---
title: "Practical ML-Project"
author: "Mrinalini"
date: "10/15/2020"
output: html_document
---


## Introduction  
Our aim in this project is to determine the manner in which 6 participants did their exercise

#### Load the required packages here

### Downloading the Data
```{r, cache = T}
trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "./data/pml-training.csv"
testFile  <- "./data/pml-testing.csv"
if (!file.exists("./data")) {
  dir.create("./data")
}
if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile=trainFile, method="curl")
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile=testFile, method="curl")
}
```  
### Read the Data
Store the csv files in two dataframes.  
```{r, cache = T}
trainRaw <- read.csv("./data/pml-training.csv")
testRaw <- read.csv("./data/pml-testing.csv")
dim(trainRaw)
dim(testRaw)
```

### Cleaning the data
making our data easier to work with
```{r, cache = T}
sum(complete.cases(trainRaw))
```
```{r, cache = T}
trainRaw <- trainRaw[, colSums(is.na(trainRaw)) == 0] 
testRaw <- testRaw[, colSums(is.na(testRaw)) == 0] 
```  
We want to get rid of the columns that do not affect the accelerometer measurements.
```{r, cache = T}
classe <- trainRaw$classe
trainRemove <- grepl("^X|timestamp|window", names(trainRaw))
trainRaw <- trainRaw[, !trainRemove]
trainCleaned <- trainRaw[, sapply(trainRaw, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testRaw))
testRaw <- testRaw[, !testRemove]
testCleaned <- testRaw[, sapply(testRaw, is.numeric)]
```

### Data Slicing
Now we split the clean training data set into a pure training data set (70%) and a validation data set (30%). This validation data set will be used to conduct cross validation in future steps.  
```{r, cache = T}
set.seed(22519) # For reproducibile purpose
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```

## Data Modeling
Using **Random Forest** algorithm, fit a predictive model for activity recognition as it automatically selects important variables. Use the **5-fold cross validation** while applying the algorithm.  
```{r, cache = T}
controlRf <- trainControl(method="cv", 5)
modelRf <- train(classe ~ ., data=trainData, method="rf", trControl=controlRf, ntree=250)
modelRf
```
Estimate the performance of the model on the validation data set.  
```{r, cache = T}
predictRf <- predict(modelRf, testData)
confusionMatrix(testData$classe, predictRf)
```
```{r, cache = T}
accuracy <- postResample(predictRf, testData$classe)
accuracy
oose <- 1 - as.numeric(confusionMatrix(testData$classe, predictRf)$overall[1])
oose
```
We see taht the estimated accuracy of the model is 99.42% and the estimated out-of-sample error is 0.58%.

## Predicting for Test Data Set
After removing the `problem_id` column, apply the model to the original testing data set.  
```{r, cache = T}
result <- predict(modelRf, testCleaned[, -length(names(testCleaned))])
result
```  

## Appendix: Figures
1. Correlation Matrix Visualization  
```{r, cache = T}
corrPlot <- cor(trainData[, -length(names(trainData))])
corrplot(corrPlot, method="color")
```
2. Decision Tree Visualization
```{r, cache = T}
treeModel <- rpart(classe ~ ., data=trainData, method="class")
prp(treeModel) # fast plot
```

## Plotting the model

```{r, cache = T}
plot(modelRf)
```
